---
title: "Downloading human liver data"
output: 
  html_document:
    toc: true
    toc_depth: 5
---

```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, cache = FALSE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)
```

```{r}
library(tidyverse)
library(SingleCellExperiment)
library(scater)
library(scran)
library(Matrix)
library(Seurat)

library(cellassign)
library(cellassign.utils)
library(scrna.utils)
```

# Data acquisition

## Normalized data

To obtain the normalized data used by the authors:

```{r, eval = FALSE}
# devtools::install_github("BaderLab/HumanLiver")
library(HumanLiver)

load(system.file("liver/HumanLiver.RData", package = "HumanLiver"))

# Put in a dummy matrix for raw.data to prevent Seurat::Convert from failing
HumanLiverSeurat@raw.data <- 
  matrix(0,
         nrow = nrow(HumanLiverSeurat@data),
         ncol = ncol(HumanLiverSeurat@data))

rownames(HumanLiverSeurat@raw.data) <- rownames(HumanLiverSeurat@data)
colnames(HumanLiverSeurat@raw.data) <- colnames(HumanLiverSeurat@data)

# Remove reduced dimension object as this causes conversion errors
HumanLiverSeurat@dr <- list()

# Convert Seurat object to SingleCellExperiment
sce_HL <- as.SingleCellExperiment(HumanLiverSeurat)
```

However, this data (both the version in `HumanLiver` and on `GEO`) is missing raw counts. To reproduce the raw counts, we processed the bam files deposited in SRA by McParland et al. with CellRanger v2.1.  

```{r, eval = FALSE}
# Read matrices from CellRanger v2.1
humanliver_filtered_dirs <- 
  Sys.glob(file.path('/datadrive/projects/cellassign-paper/data/baderlab/nick_outputs', 
                     '*', 
                     'outs', 
                     'filtered_gene_bc_matrices', 
                     'GRCh38'))
names(humanliver_filtered_dirs) <- basename(humanliver_filtered_dirs)

sce <- DropletUtils::read10xCounts(humanliver_filtered_dirs)

# Include patient as a metadata column
metadata <- data.frame(Sample=humanliver_filtered_dirs) %>%
  dplyr::mutate(patient=paste0("P", 
                               str_extract(Sample, "(?<=patient)[0-9]+"), 
                               "TLH"))

colData(sce) <- colData(sce) %>%
  data.frame(check.names = FALSE, stringsAsFactors = FALSE) %>%
  dplyr::left_join(metadata) %>%
  DataFrame(check.names = FALSE, stringsAsFactors = FALSE)

sce <- sce %>%
  scater::mutate(cellname=paste(patient, 
                                str_replace_all(Barcode, "\\-", "_"), 
                                sep = "_"))
```

## Merging raw counts with normalized SingleCellExperiment

We'll substitute in the remainder of the data for the matrix, using the values from the `HumanLiver` R package on GitHub created by the authors (`BaderLab/HumanLiver`).  

```{r, eval = FALSE}
# Use gene symbols instead of Ensembl gene IDs, to imitate McParland et al.
sce_filtered <- sce %>%
  scater::filter(cellname %in% colnames(sce_HL))

gene_ids <- get_ensembl_id(rownames(sce_HL), sce_filtered)

sce_subset <- sce_filtered[gene_ids,]
colnames(sce_subset) <- sce_subset$cellname
```

```{r, eval = FALSE}
cells_intersect <- intersect(colnames(sce_HL), 
                             colnames(sce_subset))

sce_merged <- sce_HL[,cells_intersect]
rowData(sce_merged) <- rowData(sce_subset)
rownames(sce_merged) <- rownames(sce_subset)
counts(sce_merged) <- counts(sce_subset[,cells_intersect])
sce_merged$patient <- sce_subset[,cells_intersect]$patient
```

We add cell type annotations, using the `viewHumanLiver()` function to determine the authors' mapping between clusters and cell types:

```{r, eval = FALSE}
celltype_labels <- c(
  '1'='Hepatocytes',
  '2'='ab T cells',
  '3'='Hepatocytes',
  '4'='Macrophages',
  '5'='Hepatocytes',
  '6'='Hepatocytes',
  '7'='Plasma cells',
  '8'='NK cells',
  '9'='gd T cells',
  '10'='Macrophages',
  '11'='LSECs',
  '12'='LSECs',
  '13'='LSECs',
  '14'='Hepatocytes',
  '15'='Hepatocytes',
  '16'='Mature B cells',
  '17'='Cholangiocytes',
  '18'='gd T cells',
  '19'='Erythroid cells',
  '20'='Hepatic Stellate Cells'
)

sce_merged <- sce_merged %>%
  scater::mutate(
    celltype=celltype_labels[as.character(ident)],
    patient=str_extract(patient, "[0-9]+")
  )
```

```{r}
saveRDS(sce_merged, '/datadrive/data/external/scrna/mcparland2018/human_liver.rds')
```
