---
title: "Parameter inference"
output: html_notebook
---

```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)
```

```{r, echo = FALSE}
library(knitr)
library(tidyverse)
library(tensorflow)
library(SingleCellExperiment)
library(scater)
library(data.table)
library(pheatmap)
library(ggrepel)
library(scran)
library(BiocParallel)
library(limma)
library(edgeR)
library(GSEABase)
library(scvis)
library(future)

library(scrna.utils)
library(scrna.sceutils)
library(cellassign)
library(cellassign.utils)

environment(read10xResults2) <- asNamespace("scater")

plan(multiprocess)
```

# Parameter inference

We do parameter inference with `X` set to NULL (i.e. no batch covariates). 

```{r, echo = FALSE}
set.seed(1001)

G <- 30
C <- 8
marker_prob <- 1/4
N <- 2000
P <- 1
B <- 20

rho <- matrix(rbernoulli(n = G * C, marker_prob),
              nrow = G,
              ncol = C,
              byrow = TRUE) + 0

delta <- matrix(runif(n = G * C, 
                      min = log(2),
                      max = 10),
                nrow = G,
                ncol = C,
                byrow = TRUE) * rho

beta <- matrix(rnorm(n = G * P,
                     mean = -5,
                     sd = 1),
               nrow = G,
               ncol = P,
               byrow = TRUE)

s <- rlnorm(n = N, meanlog = 0, sdlog = 1)
s <- s/mean(s)

pi <- rmultinom(n = N, size = 1, prob = rep(1/C, C)) %>% 
  apply(2, function(x) which(x == 1))

X <- matrix(1, nrow = N, ncol = P)


a <- runif(n = B, min = 0, max = 10)

Y <- simulate_cellassign(rho,
                         s,
                         pi,
                         delta,
                         B = B, 
                         a = a,
                         beta,
                         X = X,
                         min_Y = 0,
                         max_Y = 2000)
```

### Applying CellAssign

```{r, echo = FALSE}
res <- cellassign_em(exprs_obj = Y, s = s, rho = rho, X = X, B = B, use_priors = TRUE, prior_type = "shrinkage", delta_variance_prior = TRUE, verbose = FALSE)
```

## Inferred parameters

### Delta

```{r, echo = FALSE}
delta_df <- data.frame(true=as.numeric(delta), inferred=as.numeric(res$mle_params$delta))
delta_df <- delta_df %>% dplyr::filter(true != 0)

ggplot(delta_df, aes(x=true, y = inferred)) + geom_point() + theme_bw() + theme_Publication() + xlab(expression("True "~delta)) + ylab(expression("Inferred "~delta)) + geom_abline(slope = 1)
```

## Beta

```{r, echo = FALSE}
beta_df <- data.frame(true=as.numeric(beta), inferred=as.numeric(res$mle_params$beta))

ggplot(beta_df, aes(x=true, y = inferred)) + geom_point() + theme_bw() + theme_Publication() + xlab(expression("True "~beta)) + ylab(expression("Inferred "~beta)) + geom_abline(slope = 1)
```

## Cell type assignments

```{r, echo = FALSE}
table(pi, res$cell_type)
```

## Alpha (spline)

Can't be compared, since the min_Y and max_Y used in simulation and in inference are different. 


# Summary

As we can see, the model infers $\delta$ and $\beta$ values that are consistent with simulation. 

