---
title: "Follicular: FL1018"
output: html_notebook
---


```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)

knitr::opts_template$set(evalfig = list(fig.height = 6, fig.width = 7))
```

```{r, echo = FALSE}
library(knitr)
library(tidyverse)
library(tensorflow)
library(SingleCellExperiment)
library(scater)
library(data.table)
library(pheatmap)
library(ggrepel)
library(grid)
library(gage)

library(scrna.utils)
library(scrna.sceutils)
library(cellassign)
library(cellassign.utils)
library(vdj.utils)
library(Seurat)

library(infotheo)
library(lmerTest)
```

# Introduction


## Outline

* Experimental substrates
* QC stats
* Preliminary analysis
* Next steps

# Data

```{r, echo = FALSE, eval = FALSE}
data_dir <- "/datadrive/data/follicular"
filtered_matrix_dirs <- Sys.glob(file.path(data_dir, "*", "filtered_gene_bc_matrices", "GRCh38"))

names(filtered_matrix_dirs) <- stringr::str_extract(filtered_matrix_dirs, "(?<=follicular\\/).*(?=\\/filtered_gene_bc_matrices)")

filtered_matrix_dirs <- filtered_matrix_dirs[c("FL1018T1", "FL1018T2")]
```

```{r, echo = FALSE, eval = FALSE}
sce <- DropletUtils::read10xCounts(unname(filtered_matrix_dirs))
sce <- sce %>% scater::mutate(
  dataset = str_extract(Sample, "FL[0-9]+[A-Z0-9_]+"),
  sample_barcode = paste(dataset, Barcode, sep = "_")
)

sce$patient <- str_extract(sce$dataset, "^FL[0-9]+")
sce$timepoint <- str_extract(sce$dataset, "T[1-2](?=(_R[0-9]+)?$)")
sce$replicate <- str_extract(sce$dataset, "R[0-9]+$")
sce$replicate[is.na(sce$replicate)] <- "R1"
sce$transformed_progressed <- ifelse(str_detect(sce$patient, "^FL1"), "Transformed", "Progressed")
sce$status <- with(colData(sce), ifelse(timepoint == "T2", transformed_progressed, "Primary"))
```


```{r, echo = FALSE}
get_ensembl_id <- function(x, rowdat, symbol_name = "Symbol", gene_name = "ID") {
  rowdat <- as.data.frame(rowdat)
  df_as_map(rowdat %>% subset(x %in% rowdat[,symbol_name]), x, from = symbol_name, to = gene_name)
}

mito_genes <- as.character(rowData(sce)$Symbol[str_detect(rowData(sce)$Symbol, "^MT\\-")]) %>% 
  get_ensembl_id(rowData(sce))

ribo_genes <- as.character(rowData(sce)$Symbol[str_detect(rowData(sce)$Symbol, "^RP(L|S)")]) %>%
  get_ensembl_id(rowData(sce))

sce <- calculateQCMetrics(sce, exprs_values = "counts", feature_controls =
                            list(mitochondrial=mito_genes, ribosomal=ribo_genes))
```

## Mitochondrial counts

```{r, echo = FALSE}
df <- with(colData(sce), data.frame(pct_mitochondrial=pct_counts_mitochondrial,
                                    pct_ribosomal=pct_counts_ribosomal,
                                    set_id=dataset))

mito_thres <- 10

ggplot(df, aes(x=pct_mitochondrial)) + geom_histogram(aes(fill=set_id)) + theme_bw() + theme_Publication() + scale_x_continuous(breaks = log_scale_breaks(), labels = log_scale_labels(), trans = "log") + geom_vline(aes(xintercept=mito_thres), colour = "red") + xlab("% mito")
```

## Ribosomal counts

```{r, echo = FALSE}
ribo_thres <- 60

ggplot(df, aes(x=pct_ribosomal)) + geom_histogram(aes(fill=set_id)) + theme_bw() + theme_Publication() + scale_x_continuous(breaks = log_scale_breaks(), labels = log_scale_labels(), trans = "log") + geom_vline(aes(xintercept=ribo_thres), colour = "red") + xlab("% ribo")
```

## Filtering

```{r, echo = FALSE}
sce_filtered <- filter_cells(sce, nmads = 3, type = "lower", log = TRUE, max_mito = mito_thres, max_ribo = ribo_thres)
```

## Normalization

```{r, echo = FALSE}
qclust <- quickCluster(sce_filtered, min.size = 30)
sce_filtered <- computeSumFactors(sce_filtered, clusters = qclust)

sce_filtered$size_factor <- sizeFactors(sce_filtered)
```

```{r, echo = FALSE, eval = FALSE}
sce_normalized <- normalise(sce_filtered)

sce_normalized <- runPCA(sce_normalized, ntop = 1000, ncomponents = 50, exprs_values = "logcounts")
sce_normalized <- runTSNE(sce_normalized, use_dimred = "PCA", n_dimred = 50, ncomponents = 2)
sce_normalized <- runUMAP(sce_normalized, use_dimred = "PCA", n_dimred = 50, ncomponents = 2)
```

```{r, echo = FALSE}
plotTSNE(sce_normalized, colour_by = "dataset", ncomponents = 2)
```

```{r, echo = FALSE}
plotUMAP(sce_normalized, colour_by = "dataset", ncomponents = 2)
```

## Total number of UMIs

```{r, echo = FALSE}
plotPCA(sce_normalized, colour_by = "log10_total_counts")
```

## Cell cycle

```{r, echo = FALSE, eval = FALSE}
hs.pairs <- readRDS(system.file("exdata", "human_cycle_markers.rds", package="scran"))

assignments <- cyclone(sce_normalized, hs.pairs, gene.names=rowData(sce_normalized)$ID, min.iter = 10, iter = 100, verbose = TRUE, BPPARAM = MulticoreParam(15))

sce_normalized@colData <- bind_cols(sce_normalized@colData %>% as.data.frame, assignments$normalized.scores) %>% DataFrame
sce_normalized <- scater::mutate(sce_normalized, Cell_Cycle = assignments$phases)
```


```{r, echo = FALSE}
plotPCA(sce_normalized, colour_by = "Cell_Cycle")
```

```{r, echo = FALSE}
plotTSNE(sce_normalized, colour_by = "Cell_Cycle")
```

```{r, echo = FALSE, results = 'asis'}
knitr::kable(t(with(colData(sce_normalized), table(Cell_Cycle, dataset))))
```

# CellAssign

## Conservative

```{r, echo = FALSE}
follicular_markers1 <- list(
  'B cells'=c("CD19", "MS4A1", "CD79A", "CD79B", "IGKC", "IGLC2", "IGLC3", "CD74"),
  'T cells'=c("CD3D", "CD3E", "CD3G", "CD28", "TRAC")
)

follicular_rho1 <- marker_list_to_mat(follicular_markers1, include_other = FALSE)
```


```{r, echo = FALSE, results = 'asis'}
knitr::kable(follicular_rho1)
```

```{r, echo = FALSE}
B <- 40
s <- sizeFactors(sce_normalized)
sce <- sce_normalized[get_ensembl_id(rownames(follicular_rho1), rowData(sce_normalized)),]
rownames(sce) <- rownames(follicular_rho1)
counts(sce) <- as.matrix(counts(sce))

cellassign_res <- cellassign_em(exprs_obj = sce, s = s, rho = follicular_rho1, X = NULL, B = B, use_priors = TRUE, prior_type = "shrinkage", delta_variance_prior = FALSE, verbose = FALSE, em_convergence_thres = 1e-5, num_runs = 1, min_delta = log(2))
```

```{r, echo = FALSE, eval = FALSE}
sce_normalized_labeled <- sce_normalized

sce_normalized_labeled$cellassign_cluster <- factor(cellassign_res$cell_type)
sce_normalized_labeled@colData <- bind_cols(sce_normalized_labeled@colData %>% as.data.frame, as.data.frame(cellassign_res$mle_params$gamma)) %>% DataFrame(check.names = FALSE)
```

```{r, echo = FALSE, out.width = '75%'}
plotPCA(sce_normalized_labeled, ncomponents = 2, colour_by = "cellassign_cluster")
```

```{r, echo = FALSE, out.width = '75%'}
plotTSNE(sce_normalized_labeled, ncomponents = 2, colour_by = "cellassign_cluster")
```

```{r, echo = FALSE, results = 'asis'}
knitr::kable(t(table(sce_normalized_labeled$cellassign_cluster)))
```

### T-cell subclustering

```{r, echo = FALSE}
celltype_prob_threshold <- 0.9

sce_t_cells <- sce_normalized_labeled %>% 
  scater::filter(T.cells >= celltype_prob_threshold)

sce_t_cells <- runPCA(sce_t_cells, ntop = 500, ncomponents = 50)
sce_t_cells <- runTSNE(sce_t_cells, ncomponents = 2, use_dimred = "PCA", n_dimred = 50)
sce_t_cells <- runUMAP(sce_t_cells, ncomponents = 2, use_dimred = "PCA", n_dimred = 50)
```

```{r, echo = FALSE, eval = FALSE}
sce_t_cells <- cluster_wrapper(sce_t_cells, gene_subset = NULL, dimreduce_method = "PCA", clustering_method = "phenograph")
```

```{r, echo = FALSE, out.width = '75%'}
plotPCA(sce_t_cells, colour_by = "dataset")
```

```{r, echo = FALSE, out.width = '75%'}
plotTSNE(sce_t_cells, colour_by = "cluster")
```

```{r, echo = FALSE, out.width = '75%'}
plotUMAP(sce_t_cells, colour_by = "cluster")
```

```{r, echo = FALSE}
marker_tables <- findMarkers(sce_t_cells, clusters = sce_t_cells$cluster, direction = "up")

marker_tables <- lapply(marker_tables, function(x) {
  x <- x %>% 
    as.data.frame %>%
    tibble::rownames_to_column(var = "ID") %>%
    dplyr::mutate(Symbol=df_as_map(rowData(sce_t_cells) %>% as.data.frame, ID, from = "ID", to = "Symbol"))
  return(x)
})
```

## Full

```{r, echo = FALSE}
follicular_markers2 <- list(
  'B cells'=c("CD19", "MS4A1", "CD79A", "CD79B", "IGKC", "IGLC2", "IGLC3", "CD74", "SELL", "CCR7", "CXCR5"),
  'Cytotoxic T cells (activated)'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC", "CD8A", "CD8B", "GZMA", "NKG7", "CCL5", "EOMES", "IFNG", "CD69"),
  'Cytotoxic T cells'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC", "CD8A", "CD8B", "GZMA", "NKG7", "CCL5", "EOMES"),
  'Naive/resting CD4'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC", "CD4", "SELL", "CCR7", "IL7R"),
  'CD4 (activated)'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC", "CD4", "IL7R", "CD69"),
  'Tfh'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC", "CD4", "CXCR5", "PDCD1", "TNFRSF4", "ST8SIA1", "ICA1", "ICOS", "CD69") # ST8SIA1 and ICA1 very specific according to LM22
  #'T_other'=c("CD2", "CD3D", "CD3E", "CD3G", "TRAC")
  # T gamma delta -- TRDC
  # KLRB1-expressing cluster (CD4 memory T cells?? check literature) ## KLRB1 == CD161
  # KLRB1-expressing CCL5 and CD8 cluster?
  # CD8 activated -- CD69 and IFNG
) # CD28 in T cells -- but DE'd between different subsets
# what does the CD69 mean?

# P2RX5 in Tfh and B cells

follicular_rho2 <- marker_list_to_mat(follicular_markers2, include_other = TRUE)
```


```{r, echo = FALSE, results = 'asis'}
knitr::kable(follicular_rho2)
```

```{r, echo = FALSE}
B <- 40
s <- sizeFactors(sce_normalized)
sce <- sce_normalized[get_ensembl_id(rownames(follicular_rho2), rowData(sce_normalized)),]
rownames(sce) <- rownames(follicular_rho2)
counts(sce) <- as.matrix(counts(sce))

# ADD PATIENT-SPECIFIC COMPONENT TO DESIGN MATRIX

cellassign_res_full <- cellassign_em(exprs_obj = sce, s = s, rho = follicular_rho2, X = NULL, B = B, use_priors = TRUE, prior_type = "shrinkage", delta_variance_prior = FALSE, verbose = FALSE, em_convergence_thres = 1e-5, num_runs = 1, min_delta = log(2))
```

```{r, echo = FALSE, eval = FALSE}
sce_normalized_labeled_full <- sce_normalized

sce_normalized_labeled_full$cellassign_cluster <- factor(cellassign_res_full$cell_type)
sce_normalized_labeled_full@colData <- bind_cols(sce_normalized_labeled_full@colData %>% as.data.frame, as.data.frame(cellassign_res_full$mle_params$gamma)) %>% DataFrame(check.names = FALSE)
```

```{r, echo = FALSE, out.width = '75%'}
plotTSNE(sce_normalized_labeled_full, ncomponents = 2, colour_by = "cellassign_cluster")
```

### Probabilities

```{r, echo = FALSE}
ph <- plot_gamma_heatmap(cellassign_res_full, sce_normalized_labeled_full, stratified = FALSE, n_sample = 3000)

grid.newpage()
grid.draw(ph)
```

### Match to unsupervised clusters

```{r, echo = FALSE}
cellassign_cols <- c("cellassign_cluster", colnames(follicular_rho2))
cellassign_tab <- colData(sce_normalized_labeled_full) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::select(c("sample_barcode", cellassign_cols)) %>%
  dplyr::rename(cellassign_cluster_full=cellassign_cluster)

sce_t_cells_labeled <- sce_t_cells
colData(sce_t_cells_labeled) <- colData(sce_t_cells) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::left_join(cellassign_tab, by = "sample_barcode") %>%
  DataFrame(check.names = FALSE)
```

```{r, echo = FALSE}
plotUMAP(sce_t_cells_labeled, colour_by = "cluster")
```

```{r, echo = FALSE}
plotUMAP(sce_t_cells_labeled, colour_by = "cellassign_cluster_full")
```

Why is vimentin not expressed in the Tfh cells?
Why is CXCR4 expressed in the memory + Tc1 populations? 

# Malignant vs. nonmalignant B cell identification

```{r, echo = FALSE}
sce_b_cells <- sce_follicular_normalized_labeled %>% 
  scater::filter(B.cells >= celltype_prob_threshold)

sce_b_cells <- runPCA(sce_b_cells, ntop = 1000, ncomponents = 50)
sce_b_cells <- runTSNE(sce_b_cells, ncomponents = 2, use_dimred = "PCA", n_dimred = 50)

sce_mnn_b <- sce_mnn[,sce_follicular_normalized_labeled$`B cells` >= 0.9]

sce_mnn_b <- runPCA(sce_mnn_b, ntop = 1000, ncomponents = 50, exprs_values = "mnn")
sce_mnn_b <- runTSNE(sce_mnn_b, ncomponents = 2, use_dimred = "PCA", n_dimred = 50)

reducedDim(sce_b_cells, "PCA_mnn") <- reducedDim(sce_mnn_b, "PCA")
reducedDim(sce_b_cells, "TSNE_mnn") <- reducedDim(sce_mnn_b, "TSNE")
```

```{r, echo = FALSE}
plotReducedDim(sce_b_cells, use_dimred = "TSNE_mnn", colour_by = "dataset")
```

## Patient-by-patient analysis

### FL1004

```{r, echo = FALSE}
sce_FL1004_B <- sce_b_cells %>% 
  scater::filter(patient == "FL1004")

# This does way better than running PCA + TSNE again (on the non-MNN data)
plotReducedDim(sce_FL1004_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "dataset")
```

```{r, echo = FALSE}
set.seed(101)
sce_FL1004_B <- cluster_wrapper(sce_FL1004_B, gene_subset = NULL, dimreduce_method = "PCA_mnn", clustering_method = "phenograph")

plotReducedDim(sce_FL1004_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "cluster")
```

```{r, echo = FALSE}
nonmalignant_clusters <- c("3")

FL1004_B_mapping <- colData(sce_FL1004_B) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::select(sample_barcode, cluster) %>%
  dplyr::mutate(malignant_status_manual=ifelse(cluster %in% nonmalignant_clusters, "nonmalignant", "malignant"))
```

### FL1012

```{r, echo = FALSE}
sce_FL1012_B <- sce_b_cells %>% 
  scater::filter(patient == "FL1012")

set.seed(101)
sce_FL1012_B <- cluster_wrapper(sce_FL1012_B, gene_subset = NULL, dimreduce_method = "PCA_mnn", clustering_method = "phenograph")

plotReducedDim(sce_FL1012_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "dataset")
```

```{r, echo = FALSE}
plotReducedDim(sce_FL1012_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "cluster")
```

```{r, echo = FALSE}
nonmalignant_clusters <- c("2", "7", "11")

FL1012_B_mapping <- colData(sce_FL1012_B) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::select(sample_barcode, cluster) %>%
  dplyr::mutate(malignant_status_manual=ifelse(cluster %in% nonmalignant_clusters, "nonmalignant", "malignant"))
```


### FL1018

```{r, echo = FALSE}
sce_FL1018_B <- sce_b_cells %>% 
  scater::filter(patient == "FL1018")

set.seed(101)
sce_FL1018_B <- cluster_wrapper(sce_FL1018_B, gene_subset = NULL, dimreduce_method = "PCA_mnn", clustering_method = "phenograph")

plotReducedDim(sce_FL1018_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "dataset")
```

```{r, echo = FALSE}
plotReducedDim(sce_FL1018_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "cluster")
```

```{r, echo = FALSE}
nonmalignant_clusters <- c("8")

FL1018_B_mapping <- colData(sce_FL1018_B) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::select(sample_barcode, cluster) %>%
  dplyr::mutate(malignant_status_manual=ifelse(cluster %in% nonmalignant_clusters, "nonmalignant", "malignant"))
```

### FL2001

```{r, echo = FALSE}
sce_FL2001_B <- sce_b_cells %>% 
  scater::filter(patient == "FL2001")

set.seed(101)
sce_FL2001_B <- cluster_wrapper(sce_FL2001_B, gene_subset = NULL, dimreduce_method = "PCA_mnn", clustering_method = "phenograph")

plotReducedDim(sce_FL2001_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "dataset")
```

```{r, echo = FALSE}
plotReducedDim(sce_FL2001_B, ncomponents = 2, use_dimred = "TSNE_mnn", colour_by = "cluster")
```

```{r, echo = FALSE}
nonmalignant_clusters <- c("2", "9")

FL2001_B_mapping <- colData(sce_FL2001_B) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::select(sample_barcode, cluster) %>%
  dplyr::mutate(malignant_status_manual=ifelse(cluster %in% nonmalignant_clusters, "nonmalignant", "malignant"))
```

```{r, echo = FALSE}
B_mapping <- do.call(rbind,
                     list(FL1004_B_mapping, FL1012_B_mapping, FL1018_B_mapping, FL2001_B_mapping))

colData(sce_follicular_normalized_labeled) <- colData(sce_follicular_normalized_labeled) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::left_join(B_mapping %>% dplyr::select(-c(cluster))) %>%
  DataFrame

colData(sce_follicular_normalized_labeled_full) <- colData(sce_follicular_normalized_labeled_full) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::left_join(B_mapping %>% dplyr::select(-c(cluster))) %>%
  DataFrame

colData(sce_b_cells) <- colData(sce_b_cells) %>%
  data.frame(check.names = FALSE) %>%
  dplyr::left_join(B_mapping %>% dplyr::select(-c(cluster))) %>%
  DataFrame
```

```{r, echo = FALSE}
plotReducedDim(sce_b_cells, use_dimred = "TSNE_mnn", colour_by = "malignant_status_manual")
```


# Proportions and differential expression

```{r, echo = FALSE}
compute_celltype_proportions <- function(sce, celltype_col = "cellassign_cluster") {
  celltype_counts <- as.data.frame(table(colData(sce)[,"dataset"], colData(sce)[,celltype_col])) %>%
    dplyr::rename(dataset=Var1, celltype=Var2, count=Freq) %>%
    dplyr::group_by(dataset) %>%
    dplyr::mutate(proportion=count/sum(count)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(celltype=paste0(celltype, " (scrna)"))
  
  df <- reshape2::dcast(celltype_counts, formula = dataset ~ celltype, value.var = "proportion")
  
  return(df)
}
```

```{r, echo = FALSE}
sce_follicular_normalized_labeled_full$celltype <- sce_follicular_normalized_labeled_full$cellassign_cluster
```

```{r, echo = FALSE}
scrna_prop <- compute_celltype_proportions(sce_follicular_normalized_labeled_full, celltype_col = "celltype")
```

```{r, echo = FALSE}
gage_analysis <- function(sce, comparison_type = "timepoint", filter_ribo = TRUE, filter_mito = TRUE, gene_set_path, rowdat) {
  if (comparison_type == "timepoint") {
    de_table <- findMarkers(sce, clusters = sce$timepoint, block = factor(sce$patient))
    de_table <- add_symbol(de_table, filter_mito = filter_mito, filter_ribo = filter_ribo, rowdat)
    
    df <- de_table$T2
    fcvals <- df$logFC.T1
  } else if (comparison_type == "malignant") {
    sce <- sce %>% scater::filter(malignant_status %in% c("malignant", "nonmalignant"))
    
    de_table <- findMarkers(sce, clusters = sce$malignant_status, block = factor(sce$patient))
    de_table <- add_symbol(de_table, filter_mito = filter_mito, filter_ribo = filter_ribo, rowdat)
    
    df <- de_table$malignant
    fcvals <- df$logFC.nonmalignant
  } else {
    stop("Other comparisons not implemented.")
  }
  names(fcvals) <- df$Symbol
  
  gs <- gage::readList(gene_set_path)
  
  res <- gage::gage(fcvals, gsets = gs, ref = NULL, samp = NULL)
  mean_fcvals <- sapply(gs, function(x) mean(fcvals[x], na.rm = TRUE))
  
  mean_fcval_df <- data.frame(pathway=names(mean_fcvals), mean_logFC=unname(mean_fcvals))
  
  res_g <- data.frame(subset(as.data.frame(res$greater), stat.mean > 0) %>% tibble::rownames_to_column(var = "pathway"), type = "greater")
  res_l <- data.frame(subset(as.data.frame(res$less), stat.mean < 0) %>% tibble::rownames_to_column(var = "pathway"), type = "less")
  
  res_combined <- rbind(res_g, res_l)
  
  res_combined <- plyr::join(res_combined, mean_fcval_df)
  
  return(list(genes=df, pathways=res_combined))
}

add_symbol <- function(de_result, filter_ribo = TRUE, filter_mito = TRUE, rowdat) {
  de_result <- lapply(de_result, function(x) {
    x$Symbol <- df_as_map(rowdat, rownames(x), from = "ID", to = "Symbol")
    if (filter_ribo) {
      x <- subset(x, !str_detect(Symbol, "^RP"))
    }
    if (filter_mito) {
      x <- subset(x, !str_detect(Symbol, "^MT"))
    }
    return(x)
  })
  return(de_result)
}
```

```{r, echo = FALSE}
filter_ribo <- TRUE
filter_mito <- TRUE

pathway_timepoint_tfh <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "Tfh"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))

plot_markers(pathway_timepoint_tfh$pathways %>% plyr::rename(c('pathway'='Symbol', 'q.val'='FDR', 'mean_logFC'='logFC')) %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal")
```

```{r}
pathway_timepoint_cytotoxic <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "Cytotoxic T cells"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))

plot_markers(pathway_timepoint_cytotoxic$pathways %>% plyr::rename(c('pathway'='Symbol', 'q.val'='FDR', 'mean_logFC'='logFC')) %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal")
```

```{r, echo = FALSE}
pathway_timepoint_cytotoxic$pathways %>% dplyr::arrange(-mean_logFC) %>% head
```

Upregulation of PD-1 (PD-1 blockade?) in cytotoxic T cells at T2? CHECK if this corresponds to membership in the IFNG-containing subcluster of CD8 T cells. 

```{r}
pathway_timepoint_naive <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "Naive/CM CD4"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))

plot_markers(pathway_timepoint_naive$pathways %>% plyr::rename(c('pathway'='Symbol', 'q.val'='FDR', 'mean_logFC'='logFC')) %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal")
```

```{r}
pathway_timepoint_naive$pathways %>% dplyr::filter(q.val < 0.05) %>% dplyr::arrange(-mean_logFC) %>% head(10)
```


```{r}
pathway_timepoint_malignantb <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "B cells" & malignant_status_manual == "malignant"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))

plot_markers(pathway_timepoint_malignantb$pathways %>% plyr::rename(c('pathway'='Symbol', 'q.val'='FDR', 'mean_logFC'='logFC')) %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal")
```

```{r, echo = FALSE}
plot_markers(pathway_timepoint_malignantb$genes %>% as.data.frame %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal", highlight_genes = c("TCL1A", "BCL2", "BCL6", "HLA-A", "HLA-B", "HLA-C", "FCER2", "CD74"))
```

```{r}
pathway_timepoint_malignantb$pathways %>% dplyr::filter(q.val < 0.05) %>% dplyr::arrange(-mean_logFC) %>% head(10)
```

```{r}
pathway_timepoint_nonmalignantb <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "B cells" & malignant_status_manual == "nonmalignant"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))

plot_markers(pathway_timepoint_nonmalignantb$pathways %>% plyr::rename(c('pathway'='Symbol', 'q.val'='FDR', 'mean_logFC'='logFC')) %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 5, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal")
```

```{r}
pathway_timepoint_FL1018malignantb <- gage_analysis(sce = sce_follicular_normalized_labeled_full %>% scater::filter(celltype == "B cells" & malignant_status_manual == "malignant" & patient == "FL1018"), 
                                              comparison_type = "timepoint",
                                              filter_ribo = filter_ribo,
                                              filter_mito = filter_mito,
                                              gene_set_path = "/datadrive/projects/cellassign-analysis/downloads/c2.cp.reactome.v6.2.symbols.gmt", 
                                       rowdat = rowData(sce_follicular_normalized_labeled_full))
```

```{r, echo = FALSE}
plot_markers(pathway_timepoint_FL1018malignantb$genes %>% as.data.frame %>% dplyr::mutate(is_significant = FDR <= 0.05), top_n = 10, log_q_val_threshold = NULL, nrow = 1, subtitle = "T2 vs. T1", top_direction = "equal", highlight_genes = c("TCL1A", "BCL2", "BCL6", "HLA-A", "HLA-B", "HLA-C", "FCER2", "CD74"))
```

# FACS validation

Should be run with minimally-filtered dataset (i.e. filtered at 10% mito or higher), as opposed to a 5% filter that could be used for analysis. 

## Data {#bigxtable2}

```{r, echo = FALSE}
facs_table_T <- "/datadrive/data/follicular/FACS/FL_paired T pops.csv"

facs_T <- fread(facs_table_T, sep = ",", skip = 1)
colnames(facs_T)[1] <- "dataset"

facs_T <- subset(facs_T, dataset %in% unique(sce_follicular_normalized_labeled$dataset))

column_mapping <- c(
  'LIVE/Singlets\\Count\\\\'='Singlets',
  'LIVE/Singlets/Lym\\Count\\\\'='Lymphocytes',
  'LIVE/Singlets/Lym/T\\Count\\\\'='T cells',
  'LIVE/Singlets/Lym/T/CD4\\Count\\\\'='CD4 T cells',
  'LIVE/Singlets/Lym/T/CD4/CD4+Mem\\Count\\\\'='CD4 T mem',
  'LIVE/Singlets/Lym/T/CD4/CD4+Naive\\Count\\\\'='CD4 T naive',
  'LIVE/Singlets/Lym/T/CD4/Tfh\\Count\\\\'='CD4 Tfh',
  'LIVE/Singlets/Lym/T/CD8\\Count\\\\'='CD8 T cells',
  'LIVE/Singlets/Lym/T/CD8/CD8+Mem\\Count\\\\'='CD8 T mem',
  'LIVE/Singlets/Lym/T/CD8/CD8+Naive\\Count\\\\'='CD8 T naive'
)

facs_T_labeled <- facs_T %>% dplyr::select(c("dataset", names(column_mapping))) %>% plyr::rename(column_mapping)

facs_T_labeled_prop <- facs_T_labeled %>% 
  dplyr::mutate(
    `T cells` = `T cells`/Lymphocytes,
    `CD4 T cells` = `CD4 T cells`/Lymphocytes,
    `CD4 T mem` = `CD4 T mem`/Lymphocytes,
    `CD4 T naive` = `CD4 T naive`/Lymphocytes,
    `CD4 Tfh` = `CD4 Tfh`/Lymphocytes,
    `CD8 T cells` = `CD8 T cells`/Lymphocytes,
    `CD8 T mem` = `CD8 T mem`/Lymphocytes,
    `CD8 T naive` = `CD8 T naive`/Lymphocytes
  )
```

```{r, echo = FALSE}
scrna_facs_T <- facs_T_labeled_prop %>% plyr::join(scrna_prop)
metadata <- subset(colData(sce_follicular), select = c("dataset", "timepoint", "transformed_progressed", "patient")) %>% unique %>% as.data.frame
scrna_facs_T <- scrna_facs_T %>% plyr::join(metadata)
```

## T cell proportion

```{r, echo = FALSE}
ggplot(scrna_facs_T, aes(x=`T cells`, y = `Cytotoxic T cells (scrna)` + `Naive/CM CD4 (scrna)` + `T regs (scrna)` + `Tfh (scrna)`)) + 
  geom_point(aes(colour=patient)) + 
  theme_bw() + 
  theme_Publication() + 
  geom_abline(slope = 1) + 
  xlab("T cell proportion (FACS)") + 
  ylab("T cell proportion (scrna)")
```

## CD4 T cell (out of T cell) proportion

```{r, echo = FALSE}
ggplot(scrna_facs_T, aes(x=`CD4 T cells`/`T cells`, y = (`Naive/CM CD4 (scrna)` + `T regs (scrna)` + `Tfh (scrna)`)/(`Cytotoxic T cells (scrna)` + `Naive/CM CD4 (scrna)` + `T regs (scrna)` + `Tfh (scrna)`))) + 
  geom_point(aes(colour=patient)) + 
  theme_bw() + 
  theme_Publication() + 
  geom_abline(slope = 1) + 
  xlab("CD4 T/T cell proportion (FACS)") + 
  ylab("CD4 T/T cell proportion (scrna)")
```

## Tfh out of CD4 T cell proportion

```{r, echo = FALSE}
ggplot(scrna_facs_T, aes(x=`CD4 Tfh`/`CD4 T cells`, y = `Tfh (scrna)`/(`Naive/CM CD4 (scrna)` + `T regs (scrna)` + `Tfh (scrna)`))) + 
  geom_point(aes(colour=patient)) + 
  theme_bw() + 
  theme_Publication() + 
  geom_abline(slope = 1) + 
  xlab("Tfh/CD4 T cell proportion (FACS)") + 
  ylab("Tfh/CD4 T cell proportion (scrna)")
```

Generally fairly consistent. Note that quantifying memory T cells is a problem since we don't have a good marker to distinguish naive from CM. 



