---
title: "External data summary"
output: html_notebook
---

```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)

knitr::opts_template$set(evalfig = list(fig.height = 6, fig.width = 7))
```

```{r, echo = FALSE}
library(knitr)
library(tidyverse)
library(tensorflow)
library(SingleCellExperiment)
library(scater)
library(data.table)
library(pheatmap)
library(ggrepel)
library(grid)

library(scrna.utils)
library(scrna.sceutils)
library(cellassign)
library(cellassign.utils)
library(vdj.utils)
library(Seurat)

library(ExperimentHub)
library(TabulaMurisData)
library(MultiAssayExperiment)
#library(diptest)
library(infotheo)
library(DuoClustering2018)
library(lmerTest)
```

#  Tabula Muris (2018, Nature)

```{r, echo = FALSE}
eh <- ExperimentHub()
query(eh, "TabulaMurisData")

droplet <- eh[["EH1617"]]
```

```{r, echo = FALSE}
droplet$cell_ontology_class %>% table
```

```{r, echo = FALSE}
cell_type_map <- c('T cell'='T cell', 
                   'B cell'='B cell',
                   'non-classical monocyte'='monocyte',
                   'monocyte'='monocyte',
                   'classical monocyte'='monocyte',
                   'epithelial cell'='epithelial cell',
                   'endothelial cell'='endothelial cell',
                   'dendritic cell'='dendritic cell',
                   'natural killer cell'='NK cell',
                   'fibroblast'='fibroblast',
                   'granulocyte'='granulocyte'
)

droplet <- droplet %>% 
  scater::mutate(
    celltype=plyr::mapvalues(cell_ontology_class, from = names(cell_type_map),
                             to = unname(cell_type_map))
  )

droplet_filtered <- droplet %>%
  scater::filter(celltype %in% unname(cell_type_map))
```

```{r, echo = FALSE}
norm_factors <- edgeR::calcNormFactors(as.matrix(counts(droplet_filtered)), 
                                       method = "TMM")
lib_size_factors <- colSums(as.matrix(counts(droplet_filtered)))
sizeFactors(droplet_filtered) <- norm_factors * lib_size_factors/mean(norm_factors * 
                                                                        lib_size_factors)
droplet_normalized <- normalize(droplet_filtered)
droplet_normalized <- runPCA(droplet_normalized, ntop = 500, 
                             ncomponents = 50, exprs_values = "logcounts")
droplet_normalized <- runTSNE(droplet_normalized, use_dimred = "PCA", 
                              n_dimred = 50, ncomponents = 2)
```

```{r, echo = FALSE}
plotTSNE(droplet_normalized, colour_by = "channel")
```


```{r, echo = FALSE}
plotTSNE(droplet_normalized, colour_by = "celltype")
```

```{r, echo = FALSE}
marker_list <- list(
  'T cell'=c('Cd3d', 'Cd3e'),
  'Epithelial cell'=c('Epcam'),
  
)
```


TODO: Reduce this set down to a list of basic cell TYPES.
TODO: Come up with a marker gene set off the literature. (ACTUALLY: Tabula Muris themselves use a marker gene set (step 8 in their Methods))
TODO: Test on this data.

TOTHINK: Maybe do this on a celltype-by-celltype basis like they did? i.e. https://static-content.springer.com/esm/art%3A10.1038%2Fs41586-018-0590-4/MediaObjects/41586_2018_590_MOESM4_ESM.pdf

They have marker genes in their supplementary data, and it looks like they processed each organ separately and then used marker genes to make those annotations. 

# Koh et al. (2016, Scientific Data)

## Data

Referenced in Soneson/Robinson F1000 (2018). 

```{r, echo = FALSE}
maex <- readRDS('/datadrive/data/external/scrna/conquer/raw/SRP073808.rds')

print(maex)
```

```{r, echo = FALSE}
## Rounding gene-level counts. See https://www.biostars.org/p/143458/. 
cts <- round(assays(experiments(maex)[["gene"]])[["count"]])

gene_mat <- rowData(experiments(maex)[["gene"]]) %>%
  as.data.frame %>%
  dplyr::mutate(gene = str_replace_all(gene, "\\..*", ""))
rownames(gene_mat) <- gene_mat$gene
rownames(cts) <- str_replace_all(rownames(cts), "\\..*", "")

koh <- SingleCellExperiment(assays = list(counts=cts), 
                            colData = as.data.frame(colData(maex)),
                            rowData = gene_mat)
```

```{r, echo = FALSE}
koh <- koh %>% scater::mutate(
  celltype = str_replace_all(LibraryName, "^H7(_derived_|_dreived_)?", "")
)

koh$celltype %>% unique
```

```{r, echo = FALSE}
common_celltypes <- c("hESC", "APS", "MPS", "DLL1pPXM", "ESMT", "Sclrtm", "D5CntrlDrmmtm",
                      "D2LtM") #"D3GARPpCrdcM")
koh_filtered <- koh %>% 
  scater::filter(celltype %in% common_celltypes)

get_ensembl_id <- function(x, rowdat, symbol_name = "symbol", gene_name = "gene") {
  rowdat <- as.data.frame(rowdat)
  df_as_map(rowdat %>% subset(x %in% rowdat[,symbol_name]), x, from = symbol_name, to = gene_name)
}

mito_genes <- as.character(rowData(koh_filtered)$symbol[str_detect(rowData(koh_filtered)$symbol, "^MT\\-")]) %>% 
  get_ensembl_id(rowData(koh_filtered))

ribo_genes <- as.character(rowData(koh_filtered)$symbol[str_detect(rowData(koh_filtered)$symbol, "^RP(L|S)")]) %>%
  get_ensembl_id(rowData(koh_filtered))

koh_filtered <- calculateQCMetrics(koh_filtered, exprs_values = "counts", feature_controls =
                                     list(mitochondrial=mito_genes, ribosomal=ribo_genes))

koh_filtered <- filter_cells(koh_filtered, nmads = Inf, type = "lower", log = TRUE, max_mito = 30, max_ribo = 50)

norm_factors <- edgeR::calcNormFactors(as.matrix(counts(koh_filtered)), 
                                       method = "TMM")
lib_size_factors <- colSums(as.matrix(counts(koh_filtered)))
sizeFactors(koh_filtered) <- norm_factors * lib_size_factors/mean(norm_factors * 
                                                                    lib_size_factors)
koh_normalized <- normalize(koh_filtered)
koh_normalized <- runPCA(koh_normalized, ntop = 500, 
                         ncomponents = 50, exprs_values = "logcounts")
koh_normalized <- runTSNE(koh_normalized, use_dimred = "PCA", 
                          n_dimred = 50, ncomponents = 2)
```

```{r, echo = FALSE}
plotTSNE(koh_normalized, colour_by = "celltype")
```


## Bulk RNA-seq expression

```{r, echo = FALSE}
bulkrna_de <- fread('/datadrive/data/external/scrna/koh/SD_bulkRNA_DE_annotated.csv', sep = ',') %>%
  dplyr::mutate(geneID=str_replace_all(geneID, "\\..*", ""))

common_celltype_map <- c(
  'D0 H7 hESC'='hESC', 
  'D1 APS'='APS', 
  'D1 MPS'='MPS', 
  'D2 DLL1pos PXM'="DLL1pPXM", 
  'D3 Somite'="ESMT", 
  'D6 Sclerotome'="Sclrtm", 
  'D5 Dermomyotome'="D5CntrlDrmmtm",
  'D2 LatM'="D2LtM"#, 
  #'D3 Cardiac'="D3GARPpCrdcM"
)

de_mat <- bulkrna_de %>%
  dplyr::select(geneSymbol, geneID, matches("^DE")) %>%
  dplyr::select(-contains("DLL1neg")) %>%
  dplyr::select(-contains("Cardiac"))

lfc_mat <- bulkrna_de %>%
  dplyr::select(geneSymbol, geneID, matches("^sLFC")) %>%
  dplyr::select(-contains("DLL1neg")) %>%
  dplyr::select(-contains("Cardiac"))

l_rel_expr <- plyr::rbind.fill(lapply(1:nrow(lfc_mat), function(i) {
  row <- lfc_mat[i,]
  idx <- which(str_detect(colnames(row), "^sLFC"))
  lfc_cols <- colnames(row)[idx]
  
  types <- strsplit(str_replace(lfc_cols, "^sLFC ", ""), "\\-")
  types_t <- types %>% lapply(
    function(x) return(unname(common_celltype_map[x]))
  )
  
  exprs <- rep(NA, length(common_celltypes))
  names(exprs) <- common_celltypes
  exprs["hESC"] <- 0
  j <- 0
  
  while (any(is.na(exprs))) {
    known_types <- names(which(!is.na(exprs)))
    update_idxs <- sapply(types_t, function(x) (x[2] %in% known_types) & !(x[1] %in% known_types))
    base_vals <- unname(exprs[sapply(types_t[update_idxs], function(x) x[2])])
    update_types <- sapply(types_t[update_idxs], function(x) x[1])
    
    exprs[update_types] <- base_vals + as.numeric(unname(as.data.frame(row)[,idx[update_idxs]]))
  }
  
  exprs <- exprs - min(exprs)
  
  data.frame(geneSymbol=row$geneSymbol, geneID=row$geneID, exprs %>% t)
}))
```

```{r, echo = FALSE}
expr_mat <- l_rel_expr %>% 
  dplyr::filter(geneID %in% rownames(koh_normalized)) %>%
  tibble::column_to_rownames(var = "geneID") %>%
  dplyr::select(-c(geneSymbol))

expr_threshold <- log(2)

marker_idxs <- apply(expr_mat, 1, function(x) length(which(x < expr_threshold))) >= 4

binarize <- function(x, threshold) {
  x[x <= threshold] <- -Inf
  x[x > -Inf] <- 1
  x[x == -Inf] <- 0
  return(x)
}

#mean_exprs <- logcounts(koh_normalized)[rownames(expr_mat),] %>% rowMeans

#expr_mat <- expr_mat[mean_exprs > quantile(mean_exprs, c(0.5)),]
#expr_mat <- expr_mat[]
```

```{r, echo = FALSE}
selection_method <- "maxdiff"

if (selection_method == "maxdiff") {
  maxdiffs <- apply(expr_mat, 1, function(x) max(diff(sort(x))))
  thres_vals <- apply(expr_mat, 1, function(x) sort(x)[which.max(diff(sort(x)))])
  
  expr_mat_thres <- plyr::rbind.fill(lapply(1:nrow(expr_mat), function(i) {
    expr_mat[i,] <- binarize(expr_mat[i,], thres_vals[i])
  }))
  rownames(expr_mat_thres) <- rownames(expr_mat)
  rho <- expr_mat_thres[maxdiffs >= quantile(maxdiffs, c(.8)),] %>%
    as.matrix
} else if (selection_method == "dip") {
  ## Not really pvals but switched code so too lazy to change
  pvals <- apply(expr_mat, 1, function(x) dip.test(x)$statistic)
  
  rho <- expr_mat[pvals >= quantile(pvals, c(0.9)),] %>% 
    binarize(threshold = 2) %>%
    as.matrix
}

rho <- rho[intersect(rownames(rho), rownames(koh_normalized)),]
B <- 20
s <- sizeFactors(koh_normalized)
  
res <- cellassign_em(exprs_obj = koh_normalized[rownames(rho),], s = s, rho = rho, X = NULL, B = B, use_priors = TRUE, prior_type = "shrinkage", delta_variance_prior = FALSE, verbose = FALSE, em_convergence_thres = 1e-5, num_runs = 3, min_delta = 2.04)

koh_normalized$cluster <- res$cell_type

evaluation_measures <- compute_evaluation_measures(koh_normalized, 
                                                   truth_labels = koh_normalized$celltype,
                                                   inferred_labels = koh_normalized$cluster)
```

```{r, echo = FALSE}
plotTSNE(koh_normalized, colour_by = rownames(rho)[2])
```

```{r, echo = FALSE}
plotTSNE(koh_normalized, colour_by = "celltype")
```

```{r, echo = FALSE}
plotTSNE(koh_normalized, colour_by = "cluster")
```

```{r, echo = FALSE}
koh_normalized$dataset <- koh_normalized$celltype

ph <- plot_gamma_heatmap(res, koh_normalized, stratified = FALSE, n_sample = 3000)

grid.newpage()
grid.draw(ph$gtable)
```


## Delta comparison

```{r, echo = FALSE}
rel_expr_thres <- (expr_mat - thres_vals)[maxdiffs >= quantile(maxdiffs, c(.8)),] * rho

rel_expr_df <- reshape2::melt(rel_expr_thres %>% as.matrix) %>%
  dplyr::rename(gene=Var1, celltype=Var2, relexpr=value) %>%
  dplyr::filter(relexpr != 0)

delta_df <- reshape2::melt(res$mle_params$delta) %>% 
  dplyr::rename(gene=Var1, celltype=Var2, delta=value) %>%
  dplyr::filter(delta != 0)

delta_rel_expr <- rel_expr_df %>% 
  inner_join(delta_df)
```

```{r, echo = FALSE}
mod <- lmer(delta ~ relexpr + (1|gene), delta_rel_expr)

summary(mod)

library(MuMIn)

MuMIn::r.squaredGLMM(mod)
```

TODO: Show TSNE + clusters for our method vs. Seurat and SC3. Key highlight: They keep APS and MPS merged while we are capable of separating. 
TODO: Do a delta comparison of true deltas compared to 'baseline' (maybe just define this as the highest one below the maxdiff) vs. inferred deltas. 

## Unsupervised methods

```{r, echo = FALSE}
cluster_wrapper <- function(sce, gene_subset = NULL, dimreduce_method, clustering_method, conda_env = "r-tensorflow", conda_binary = "/home/rstudio/miniconda/bin/conda", seurat_resolution = 0.6,
                            object2 = NULL, object2_cluster_label = "Group") {
  if (!is.null(gene_subset)) {
    sce <- sce[gene_subset,]
    sce <- runPCA(sce, ntop = 500, ncomponents = 50, exprs_values = "logcounts")
    
    if (!is.null(object2)) {
      object2 <- object2[gene_subset,]
    }
  }
  
  if (clustering_method == "phenograph") {
    sce_clustered <- cluster_cells(sce, method = clustering_method, dimreduce_type = dimreduce_method, conda_env = conda_env, 
      conda_binary = conda_binary,
      phenograph_module = "scrnatools.methods.clustering.phenograph_analysis")
  } else if (clustering_method == "dbscan") {
    sce_clustered <- cluster_cells(sce, method = clustering_method, dimreduce_type = dimreduce_method, dbscan_epsilon = 0.5, dbscan_minPoints = 5)
  } else if (str_detect(clustering_method, "^seurat")) {
    sce_clustered <- cluster_cells(sce, method = clustering_method, dimreduce_type = dimreduce_method,
                                   seurat_resolution = seurat_resolution)
  } else if (clustering_method %in% c("Zheng_cor", "scmap")) {
    sce_clustered <- cluster_cells(sce, method = clustering_method, dimreduce_type = dimreduce_method,
                                   object2 = object2, object2_cluster_label = object2_cluster_label)
  } else {
    sce_clustered <- cluster_cells(sce, method = clustering_method, dimreduce_type = dimreduce_method)
  }
  return(sce_clustered)
}

map_and_evaluate_clusters <- function(sce_sim, min_count = 0) {
  # Remove genes that are lowly expressed, and cells with no UMIs
  sce_sim <- sce_sim[rowSums(counts(sce_sim)) >= min_count,colSums(counts(sce_sim)) > 0]
  
  nclusts <- length(unique(sce_sim$cluster))
  
  if (nclusts > 1) {
    sce_de <- map_clusters(sce_sim, method = "de2", FDR_cutoff = 0.05)
    sce_correlation <- map_clusters(sce_sim, method = "correlation", min_correlation = 0)
  } else {
    max_group <- names(which.max(table(sce_sim$Group)))[1]
    
    sce_de <- sce %>% scater::mutate(inferred_group=max_group)
    sce_correlation <- map_clusters(sce_sim, method = "correlation", min_correlation = 0)
  }
  
  de_evaluation_measures <- compute_evaluation_measures(sce_de, 
                                                         truth_labels = sce_de$Group,
                                                         inferred_labels = sce_de$inferred_group)
  corr_evaluation_measures <- compute_evaluation_measures(sce_correlation, 
                                                           truth_labels = sce_correlation$Group,
                                                           inferred_labels = sce_correlation$inferred_group)
  
  evaluation_measures <- rbind(data.frame(mapping_type = "de", de_evaluation_measures, stringsAsFactors = FALSE), data.frame(mapping_type = "correlation", corr_evaluation_measures, stringsAsFactors = FALSE))
  
  return(evaluation_measures)
}
```

```{r, echo = FALSE}
#koh_normalized <- runPCA(koh_normalized, ncomponents = 50, ntop = 1000)
#koh_normalized <- runTSNE(koh_normalized, ncomponents = 2, use_dimred = "PCA", n_dimred = 50)

koh_normalized$Group <- koh_normalized$celltype
koh_normalized$Batch <- 1
rowData(koh_normalized)$Gene <- rownames(rowData(koh_normalized))
rowData(koh_normalized)$symbol <- rownames(rowData(koh_normalized))
clustering_method <- "seurat"
dimreduce_method <- "PCA"
markers_to_use <- rownames(rho)
seurat_resolution <- 0.8

sce_sim_full <- cluster_wrapper(koh_normalized, gene_subset = NULL, dimreduce_method = dimreduce_method, clustering_method = clustering_method, seurat_resolution = seurat_resolution)

print(with(colData(sce_sim_full), table(cluster, Group)))

sce_correlation <- map_clusters(sce_sim_full, method = "correlation", min_correlation = 0)
corr_evaluation_measures_full <- compute_evaluation_measures(sce_correlation, 
                                                        truth_labels = sce_correlation$Group,
                                                        inferred_labels = sce_correlation$inferred_group)

sce_sim_markers <- cluster_wrapper(koh_normalized, gene_subset = markers_to_use, dimreduce_method = dimreduce_method, clustering_method = clustering_method, seurat_resolution = seurat_resolution)

print(with(colData(sce_sim_markers), table(cluster, Group)))
sce_correlation <- map_clusters(sce_sim_markers, method = "correlation", min_correlation = 0)
corr_evaluation_measures_markers <- compute_evaluation_measures(sce_correlation, 
                                                        truth_labels = sce_correlation$Group,
                                                        inferred_labels = sce_correlation$inferred_group)
```

# Zheng et al. (2017, Nature Comms)

```{r}
data_dir <- "/datadrive/projects/cellassign-analysis/data/pure_pbmc"
filtered_matrix_dirs <- Sys.glob(file.path(data_dir, "*", "filtered_matrices_mex", "hg19"))

names(filtered_matrix_dirs) <- stringr::str_extract(filtered_matrix_dirs,          "(?<=pure_pbmc\\/).*(?=\\/filtered_matrices)")
```

```{r, echo = FALSE, eval = FALSE}
sce_pbmc <- DropletUtils::read10xCounts(unname(filtered_matrix_dirs))

sce_pbmc <- sce_pbmc %>%
  scater::mutate(celltype=basename(dirname(dirname(Sample))),
                 sample_barcode=paste0(Sample, "_", Barcode))
```


```{r}
mito_genes <- get_mitochondrial_genes(filtered_matrix_dirs[1], type = "gene_id", check_db = FALSE, species = "human")
ribo_genes <- as.character(rowData(sce_pbmc)$Symbol[str_detect(rowData(sce_pbmc)$Symbol, "^RP(L|S)")]) %>%
  get_ensembl_id(rowData(sce_pbmc), symbol_name = "Symbol", gene_name = "ID")

sce_pbmc <- calculateQCMetrics(sce_pbmc, exprs_values = "counts", feature_controls =
                                 list(mitochondrial=mito_genes, ribosomal=ribo_genes))

sce_pbmc_filtered <- filter_cells(sce_pbmc, nmads = 3, type = "lower", log = TRUE, max_mito = 5, max_ribo = 60)
```

```{r, echo = FALSE}
sample_df <- colData(sce_pbmc_filtered) %>%
  as.data.frame %>%
  splitstackshape::stratified(group = "celltype", size = 2000)

colnames(sce_pbmc_filtered) <- sce_pbmc_filtered$sample_barcode
sce_pbmc_sample <- sce_pbmc_filtered[,sample_df$sample_barcode]
```

```{r, echo = FALSE}
norm_factors <- edgeR::calcNormFactors(as.matrix(counts(sce_pbmc_sample)), 
                                       method = "TMM")
lib_size_factors <- colSums(as.matrix(counts(sce_pbmc_sample)))
sizeFactors(sce_pbmc_sample) <- norm_factors * lib_size_factors/mean(norm_factors * 
                                                                       lib_size_factors)
sce_pbmc_normalized <- normalize(sce_pbmc_sample)
sce_pbmc_normalized <- runPCA(sce_pbmc_normalized, ntop = 500, 
                              ncomponents = 50, exprs_values = "logcounts")
sce_pbmc_normalized <- runTSNE(sce_pbmc_normalized, use_dimred = "PCA", 
                               n_dimred = 50, ncomponents = 2)
```



# Scratch

Not using this as classes aren't specific enough. 

```{r, echo = FALSE}
sce_full_Zhengmix4eq <- sce_full_Zhengmix4eq(metadata = FALSE)

sce_full_Zhengmix4eq <- runPCA(sce_full_Zhengmix4eq, ntop = 500, 
                               ncomponents = 50, exprs_values = "logcounts")
sce_full_Zhengmix4eq <- runTSNE(sce_full_Zhengmix4eq, use_dimred = "PCA", 
                                n_dimred = 50, ncomponents = 2)

sce_full_Zhengmix4eq <- sce_full_Zhengmix4eq %>%
  scater::mutate(celltype = phenoid)
```

```{r, echo = FALSE}
plotTSNE(sce_full_Zhengmix4eq, colour_by = "celltype")
```

```{r, echo = FALSE}

```





