---
title: "Simulation result summary"
output:
  html_document:
    df_print: paged
---

```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)

knitr::opts_template$set(evalfig = list(fig.height = 6, fig.width = 7))
```

```{r, echo = FALSE}
library(knitr)
library(tidyverse)
library(tensorflow)
library(SingleCellExperiment)
library(scater)
library(data.table)
library(pheatmap)
library(ggrepel)
library(grid)

library(scrna.utils)
library(scrna.sceutils)
library(cellassign)
library(cellassign.utils)
library(vdj.utils)
```

## Functions

```{r}
load_annotation_files <- function(output_dir, pattern = "*_eval_measures.tsv") {
  param_files <- Sys.glob(file.path(output_dir, "*", "*_params.tsv"))
  ann_files <- Sys.glob(file.path(output_dir, "*", pattern))
  
  params <- plyr::rbind.fill(lapply(param_files, function(f) {
    params <- fread(f)
    data.frame(root=dirname(f), params)
  }))
  
  anns <- plyr::rbind.fill(lapply(ann_files, function(f) {
    anns <- tryCatch({
      df <- fread(f)
      return(data.frame(root=dirname(f), df))
    }, error = function(e) {
      return(data.frame())
    })
  }))
  
  result <- anns %>% plyr::join(params, by = "root")
  return(result)
}

evaluate_de <- function(eval_measures, measures = c("v_measure", "accuracy", "macro_f1", "micro_f1"),
                        x_var = "de_prob") {
  if ("mapping_type" %in% names(eval_measures)) {
    eval_measures <- eval_measures %>%
      dplyr::filter(is.na(mapping_type) | mapping_type == "de")
  }
  
  if ("gene_set" %in% names(eval_measures)) {
    eval_measures_markers <- eval_measures %>%
      dplyr::filter(is.na(gene_set) | gene_set == "markers")
    
    eval_measures_full <- eval_measures %>%
      dplyr::filter(is.na(gene_set) | gene_set == "full")
    
    elist <- list('markers'=eval_measures_markers,
                  'full'=eval_measures_full)
  } else {
    elist <- list('cellassign'=eval_measures)
  }
  
  plots <- lapply(elist, function(gs) {
    eval_measures_markers_de_melted <- gs %>%
      reshape2::melt(measure.vars = measures,
                     variable.name = "measure",
                     value.name = "value") %>%
      dplyr::mutate_(.dots = setNames(list(lazyeval::interp(~factor(x), x=as.name(x_var))), "xval"))
    
    marker_de_plot <- ggplot(eval_measures_markers_de_melted, aes(x=xval, y = value, fill = clustering_method)) + 
      geom_boxplot(outlier.size = -1) + 
      theme_bw() + 
      theme_Publication() + 
      theme_nature() + 
      facet_wrap(~ measure, scales = "free", ncol = 1, strip.position = "right") + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      xlab("% genes DE/group") + 
      ylab("Score")
    
    return(marker_de_plot)
  })
  
  return(plots)
}
```

# DE prob simulation results

## Resultsdeprob2

```{r, echo = FALSE}
output_dir <- "/datadrive/pipeline/results/cellassign-sim-comparison/resultsdeprob4"

eval_measures <- load_annotation_files(output_dir, pattern = "*_eval_measures.tsv")
delta_vals <- load_annotation_files(output_dir, pattern = "*_delta_compare.tsv")
```

```{r}
de_res <- evaluate_de(eval_measures, measures = c("v_measure",
                                                  "accuracy",
                                                  "micro_f1",
                                                  "macro_f1"),
                      x_var = "de_prob")
```

## CellAssign vs. other methods (only seeing marker genes)

```{r, opts.label='evalfig'}
de_res$markers
```

## CellAssign vs. other methods (seeing all genes)

```{r, opts.label='evalfig'}
de_res$full
```

## Resultsdeprob3

```{r, echo = FALSE}
output_dir <- "/datadrive/pipeline/results/cellassign-sim-comparison/resultsdeprob5"

eval_measures <- load_annotation_files(output_dir, pattern = "*_eval_measures.tsv")
delta_vals <- load_annotation_files(output_dir, pattern = "*_delta_compare.tsv")
```

```{r}
de_res <- evaluate_de(eval_measures, measures = c("v_measure",
                                                  "accuracy",
                                                  "micro_f1",
                                                  "macro_f1"),
                      x_var = "de_prob")
```

## CellAssign vs. other methods (only seeing marker genes)

```{r, opts.label='evalfig'}
de_res$markers
```

## CellAssign vs. other methods (seeing all genes)

```{r, opts.label='evalfig'}
de_res$full
```

# Wrong marker simulation results

## Wrongmarkers1

```{r, echo = FALSE}
output_dir <- "/datadrive/pipeline/results/cellassign-sim-comparison/wrongmarkers1"

eval_measures <- load_annotation_files(output_dir, pattern = "*_eval_measures.tsv")
delta_vals <- load_annotation_files(output_dir, pattern = "*_delta_compare.tsv")
```

## Marker genes/class: 15, scale: 0.03

```{r, echo = FALSE}
res <- evaluate_de(eval_measures %>% dplyr::filter(max_genes == 15,
                                                   de_facscale == 0.03), measures = c("v_measure",
                                               "accuracy",
                                               "micro_f1",
                                               "macro_f1"),
                   x_var = "wrong_marker_proportion")

res$cellassign
```

## Marker genes/class: 15, scale: 0.06

```{r, echo = FALSE}
res <- evaluate_de(eval_measures %>% dplyr::filter(max_genes == 15,
                                                   de_facscale == 0.06), measures = c("v_measure",
                                               "accuracy",
                                               "micro_f1",
                                               "macro_f1"),
                   x_var = "wrong_marker_proportion")

res$cellassign
```


## Wrongmarkers2

```{r, echo = FALSE}
output_dir <- "/datadrive/pipeline/results/cellassign-sim-comparison/wrongmarkers2"

eval_measures <- load_annotation_files(output_dir, pattern = "*_eval_measures.tsv")
delta_vals <- load_annotation_files(output_dir, pattern = "*_delta_compare.tsv")
```

## Marker genes/class: 5, scale: 0.03

```{r, echo = FALSE}
res <- evaluate_de(eval_measures %>% dplyr::filter(max_genes == 5,
                                                   de_facscale == 0.03), measures = c("v_measure",
                                               "accuracy",
                                               "micro_f1",
                                               "macro_f1"),
                   x_var = "wrong_marker_proportion")

res$cellassign
```

## Marker genes/class: 5, scale: 0.09

```{r, echo = FALSE}
res <- evaluate_de(eval_measures %>% dplyr::filter(max_genes == 5,
                                                   de_facscale == 0.09), measures = c("v_measure",
                                               "accuracy",
                                               "micro_f1",
                                               "macro_f1"),
                   x_var = "wrong_marker_proportion")

res$cellassign
```

